Supplementary Information: Quantifying the pathways to life using assembly spaces

### 2021-05-21 papers/assembly/I/main.tex
Annotate some painpoints regarding bases.
Petri net morphisms do not, in general, preserve bases. Given that's
case, we may have to losen our definition of an Assembly space to "a
special Petri net with chosen generating set".

            \end{pgfonlayer}
        \end{tikzpicture}
    }
\end{center}
where the generating set $\{A, AB, C\}$ is mapped into the set $\{X, Y\}$, both blue.
The set $\{X, Y\}$ generates the entire image of $P$, whose edges are red.
This is not to say that the only generating sets in the image of a Petri net will be the image of a generating set, for example $\{X\}$ is generating in the image $f(P)$ but its preimage $\{A,B\}$ is not.

Still, the fact that generating sets are preserved yields nicely our final result: bases are preserved under Petri net morphisms
\begin{corollary}\label{lem:image-basis}
    Let $P$ be a Petri net with basis $\B$ and let $f : A \rightarrow B$ be a Petri net morphism.
    Then $f(\B)$ is a basis for $f(P)$.
\end{corollary}
\begin{proof}
    \TODO{Complete}
\end{proof}
The property that Petri net morphisms preserve bases is fundamentally important for establishing relationships between basis-dependent quantities computed in one Petri net with those in another.

@@

This too will alway be the case: the image of a rooted subspace will always be rooted in the image.
\begin{theorem}\label{thm:rooted-image}
    Let $A$ be an assembly space, $Z \subseteq A$ a rooted assembly subspace, and $f : A \rightarrow B$ an assembly map.
    Then $(f(Z), S(f)(\B_Z))$ is rooted in $(f(A), S(f)(\B_A)$.
\end{theorem}
\begin{proof}
    \TODO{Complete}
\end{proof}

@@

The first task, when presented with an assembly space, is to determine the assembly index of the sites in the space.
In practical terms, this requires an algorithm for computing the assembly index relative to some space.
In this work, we do not present a general algorithm to do so, though a simple depth-first search of the space may suffice in some cases.
Rather, we first show that the assembly index is in fact computable.
\begin{theorem}\label{thm:computable}
    If $A$ is an assembly space and $x \in S(A)$ is finite, then $c(x)$ is computable.
\end{theorem}
\begin{proof}
    \TODO{Complete}
\end{proof}
Of course, just because the value is computable does not mean that it can be computed in reasonable time.
In fact, the general problem is likely to be \texttt{NP}-hard.
In situtations where direct computation is infeasible, we can resort to bounding the value of the assembly index by computing the index in an analogous space.
The first of the bounds we present is an upper bound determined by computing the assembly index in a subspace containing the site.
\begin{theorem}\label{cor:subspace-index-upper-bound}
    Let $(A, \B_A)$ be an assembly space and let $(B, \B_A)$ be a rooted assembly subspace of $A$.
    For every finite $b \in S(B)$, the assembly index of $b$ in $B$ is at least the assembly index of $b$ in $A$.
    That is, $c_B(b) \ge c_A(b)$ for all $b \in S(B)$
\end{theorem}
\begin{proof}
    \TODO{Complete}
\end{proof}

@@

The second bound we establish is a lower bound determined by mapping the space in question into a space within which computing the assembly index may be simpler.
\begin{theorem}\label{thm:image-lower-bound}
    If $f: A \rightarrow B$ is an assembly map, then $c_{f(A)}(f(x)) \le c_A(x)$ for all finite $x \in S(A)$.
    Furthermore, if $f$ is an embedding, then $c_{f(A)}(f(x)) = c_A(x)$.
\end{theorem}
\begin{proof}
    \TODO{Complete}
\end{proof}

>>>>

            \end{pgfonlayer}
        \end{tikzpicture}
    }
\end{center}
where the generating set $\{A, AB, C\}$ is mapped into the set $\{X, Y\}$, both blue.
The set $\{X, Y\}$ generates the entire image of $P$, whose edges are red.
This is not to say that the only generating sets in the image of a Petri net will be the image of a generating set, for example $\{X\}$ is generating in the image $f(P)$ but its preimage $\{A,B\}$ is not. The property that Petri net morphisms preserve generating sets is fundamentally important for establishing relationships between generator-dependent quantities computed in one Petri net with those in another.

@@

This too will always be the case: the image of a rooted subspace will always be rooted in the image.
\begin{theorem}\label{thm:rooted-image}
    Let $A$ be an assembly space, $Z \subseteq A$ a rooted assembly subspace, and $f : A \rightarrow B$ an assembly map.
    Then $(f(Z), S(f)(\B_Z))$ is rooted in $(f(A), S(f)(\B_A)$.

    \TODO{This almost certainly isn't true as defined at present. Assembly maps do not, in general, preserve the basis, $S(f)(\B_A)$ and $S(f)(\B_Z)$ are probably not bases. That said, we might not actually need this. It'll depend on how the proof of \Cref{thm:image-lower-bound} go.}
\end{theorem}
\begin{proof}
    \TODO{Complete}
\end{proof}

@@

The first task, when presented with an assembly space, is to determine the assembly index of the sites in the space.
In practical terms, this requires an algorithm for computing the assembly index relative to some space.
In this work, we do not present a general algorithm to do so, though a simple depth-first search of the space may suffice in some cases.
Rather, we first show that the assembly index is in fact computable.
\begin{theorem}\label{thm:computable}
    If $A$ is an assembly space and $x \in S(A)$ is finite, then $c(x)$ is computable.
\end{theorem}
\begin{proof}
    \TODO{Complete}
\end{proof}
Of course, just because the value is computable does not mean that it can be computed in reasonable time.
In fact, the general problem is likely to be \texttt{NP}-hard.
In situtations where direct computation is infeasible, we can resort to bounding the value of the assembly index by computing the index in an analogous space.
The first of the bounds we present is an upper bound determined by computing the assembly index in a subspace containing the site.
\begin{theorem}\label{thm:subspace-index-upper-bound}
    Let $(A, \B_A)$ be an assembly space and let $(B, \B_A)$ be a rooted assembly subspace of $A$.
    For every finite $b \in S(B)$, the assembly index of $b$ in $B$ is at least the assembly index of $b$ in $A$.
    That is, $c_B(b) \ge c_A(b)$ for all $b \in S(B)$
\end{theorem}
\begin{proof}
    \TODO{Complete}
\end{proof}

@@

The second bound we establish is a lower bound determined by mapping the space in question into a space within which computing the assembly index may be simpler.
\begin{theorem}\label{thm:image-lower-bound}
    If $f: A \rightarrow B$ is an assembly map, then $c_{f(A)}(f(x)) \le c_A(x)$ for all finite $x \in S(A)$.
    Furthermore, if $f$ is an embedding, then $c_{f(A)}(f(x)) = c_A(x)$.
\end{theorem}
\begin{proof}
    \TODO{Complete}

    \red{This may suffer from the fact that assembly maps do not preserve bases. If that's the case, we should rip all mention of bases out for a future work.}
\end{proof}
